Playing Games: A Case Study in Active Learning Applied to Game Theory

### 2015-06-21 paper/paper.tex
Initial description of two thirds game done.

\documentclass[a4paper]{article}

@@

A variety of studies have highlighted the effectiveness of active learning
\cite{Freeman2014, Hake1998, Prince2004}. These two papers are in fact meta studies
evaluating the effectiveness an active student centred approach. Note that the
definition used in \cite{Freeman2014} corresponds to simply any pedagogic
approach in which students are not passive consumers of a lecture during the
class meeting.

Some examples of active learning in a variety of subjects include:

@@

The above sources (and references therein) generally discuss the pedagogic
approach from a macroscopic point of view with regards to the course considered.
This manuscript will give a detailed description of two particular active
learning activities used in the instruction of Game Theoretic concepts.

\section{An exemplar: a course in game theory}

\begin{itemize}
    \item Discuss other examples of how game theory is taught
    \item Describe this course
    \item Give description of the general philosophy of the course (no content:
        purely examples, which give data for games)
\end{itemize}

\subsection{Best responses}

\begin{itemize}
    \item The two thirds of the average game: how this is a basic game
    \item How it is played in class
    \item Description of data and discussion that ensues
\end{itemize}

\subsection{Repeated and random games}

>>>>

\documentclass{article}

@@

A variety of studies have highlighted the effectiveness of active learning
\cite{Freeman2014, Hake1998, Prince2004}. These two papers are in fact meta studies
evaluating the effectiveness an active student centred approach. Note that the
definition used in \cite{Freeman2014} corresponds to simply any pedagogic
approach in which students are not passive consumers of a lecture during the
class meeting. Some examples of active learning in a variety of subjects include:

@@

The above sources (and references therein) generally discuss the pedagogic
approach from a macroscopic point of view with regards to the course considered.
This manuscript will give a detailed description of two particular active
learning activities used in the instruction of Game Theoretic concepts:

\begin{itemize}
    \item Section \ref{sec:best_responses} will describe an in class activity
        and software package used to introduce students to the topic of best
        response dynamic \cite{Maschler2013}.
    \item Section \ref{sec:repeated_games} will describe an implementation of
        Axelrod's tournament \cite{Axelrod1980a,Axelrod1980a}.
\end{itemize}

These activities aim to introduce the student to the concepts and aspire to
their curiosity as to the underlying mathematics. Note that if there is any
doubt as to the effectiveness of active learning approaches, for example this paper
(the only one that this author could identify) \cite{Andrews2011} identifies no
such relationship are still beneficial to the students' learning.
Indeed in \cite{Poropat2014}  the greatest predictors of
academic performance are identified not as general intelligence \cite{Wright1905} but
personality factors such as conscientiousness and openness.

\section{An exemplar: a course in game theory}\label{sec:game_theory}

Game Theory as a topic is well suited to approaches that use activities
involving students as players to introduce the concepts, rules and strategies
for particular games and/or theorems presented.

In \cite{Brokaw2004} one such activity is presented: a game that allow players
to grasp the concept of common knowledge of rationality. Another good example is
\cite{Polak2008}: Yale's Professor Polak's course, the videos available at that
reference (a YouTube playlist) all show that students are introduced to every
concept through activity before discussing theory.

Just as the activity presented in \cite{Brokaw2004} the activities presented
here are both suited for as an early introduction to the concepts (although the
activity of Section \ref{sec:repeated_games} is potentially better suited to
being used at a later stage). Furthermore, these activities have also been used
as outreach activities for high school students with no knowledge of further
mathematics.

\subsection{Best response dynamics}\label{sec:best_responses}

The first step in this activity and potentially before any prior description of
Game Theory students are invited to answer the following simple question:

\begin{center}
    \textbf{What is a game?}
\end{center}

Through discussion the class will usually arrive at the following consensus:

\begin{itemize}
    \item A game must have a certain number \(N\geq 1\) of players;
    \item Each player must have available to them a certain number of strategies
        that define what they can do;
    \item Once all players have chosen their strategy, rules must specify what
        the outcome is.
\end{itemize}

This corresponds to the general definition of a strategic form game
\cite{Maschler2013}. The main goal of this activity is to not only understand
the vocabulary but also the important concept of response dynamics which aims
to identify what is the best option given prior knowledge of all other players
\cite{Maschler2013}. One particular game that can be analysed using base
response dynamics is often referred to:

\begin{center}
    \textbf{The two thirds of the average game.}
\end{center}

A good description of the game and the human dynamics associated to the play is
given in \cite{Nagel1995}. The rules are as follows:

\begin{itemize}
    \item All players choose a number between 0 and 100;
    \item The player whose choice was closest to \(\frac{2}{3}\) of the average
        of the choices wins.
\end{itemize}

The activity is similar to the activity described in
\cite{TheEconomicsNetwork2013}:

\begin{enumerate}
    \item Students are handed out copies of the form available at URL
    \item Students are explained the rules of the game and invited to play a
        first guess.
    \item After this the class is explained the rationalisation of the game
        (which uses best response dynamics to iteratively eliminate dominated
        strategies) as shown in Figure
        \ref{fig:rationalisation_of_two_thirds_game}.
    \item Students are invited to play one more time.
    \item Results are analysed and presented to the class for discussion.
\end{enumerate}

\subsection{Repeated and random games}\label{sec:repeated_games}
